---
title: "Mosaic and the Less Volume, More Creativity Approach"
author: "Nicholas Horton (nhorton@amherst.edu)"
date: "August 2, 2020"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r, setup, include=FALSE}
require(mosaic)   # Load additional packages here 
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 4)
```

# Introduction

This is an illustrated example of the analyses presented in Nick Horton's JSM 2020 talk "The role of computing at the core of a modern introductory statistics course".  See https://github.com/Amherst-Statistics/jse for slides and https://cran.r-project.org/web/packages/mosaic/vignettes/mosaic-resources.html for more information about resources related to mosaic.


## Preliminary code
```{r}
library(mosaic)
library(palmerpenguins)

twoisland <- penguins %>%
  mutate(island = as.character(island)) %>%
  filter(island != "Torgersen")
```

It's puzzling to me that `palmerpenguins` has island as a factor.  This complicates dropping one of the islands so that we can easily demonstrate using a two-sample t-test.  (I would generally hide this code from students in the first few weeks.)

# Means of two groups

### Mosaic code

```{r}
# this should work in base R, but alas, it doesn't
mosaic::mean(
  bill_length_mm ~ island, 
  na.rm = TRUE, 
  data = twoisland)

df_stats(
  bill_length_mm ~ island, 
  data = twoisland) # mosaic helper function
```

One command provides a set of summaries (and provides sample size and missing values).

### Equivalent in the tidyverse

```{r message = FALSE}
twoisland %>%
  group_by(island) %>%
  summarize(
    billmean = mean(
      bill_length_mm, 
      na.rm = TRUE), 
    n = n()
  )
```

# Graphical displays

### Mosaic approach (using ggformula)

```{r warning = FALSE}
gf_boxplot(
  bill_length_mm ~ island, 
  data = twoisland)
```

The ggformula package provides a formula interface to ggplot2 graphics

### Equivalent in the tidyverse

```{r warning = FALSE}
ggplot(
  twoisland, 
  aes(
    y = bill_length_mm, 
    x = island)) +
  geom_boxplot()
```

While ggplot2 is very powerful, some aspects of the syntax (`aes()` and `+`) do not translate from the equivalent comments for descriptive statistics and modeling.

\clearpage

# Two sample t-test 

### base R
```{r}
t.test(
  bill_length_mm ~ island, 
  var.equal = TRUE, 
  data = twoisland)
```

\clearpage

### another approach in base R

```{r}
library(infer)
modttest <- lm(
  bill_length_mm ~ island, 
  data = twoisland)
confint(modttest)
msummary(modttest)

```

# Multiple regression

The ggformula package can be used to general scatterplots colored by species.  Note that in this usage the same *pipe* operator (`%>%`) is used as in the tidyverse to add the regression lines for each group.

```{r warning = FALSE}
gf_point(
  bill_length_mm ~ bill_depth_mm, 
  color = ~ species, 
  data = twoisland) %>%
  gf_lm()
modmultreg <- lm(
  bill_length_mm ~ bill_depth_mm + species, 
  data = twoisland)
msummary(modmultreg)
confint(modmultreg)
```

\clearpage

# Bootstrapping a multiple regression model


```{r bootstrap, cache = TRUE}
set.seed(1619)
bootstraps <- do(5000) * 
  lm(
    bill_length_mm ~ bill_depth_mm + species, 
    data = resample(twoisland))
```

```{r}
qdata(
  ~ speciesGentoo, 
  p = c(.025, .975), 
  data = bootstraps)
```





